---
title: "Creating Kaplan-Meier Survival Plots with tflmetaR and gridify"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating Kaplan-Meier Survival Plots with tflmetaR and gridify}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
```

## Introduction

This vignette demonstrates how to create publication-quality Kaplan-Meier survival plots using the `survminer` package for visualization, `gridify` for professional annotation layouts, and `tflmetaR` for centralized metadata management.

The workflow covers:

- Loading and preparing ADTTE (Time-to-Event) data from CDISC-compliant datasets
- Fitting survival models and generating Kaplan-Meier curves
- Combining survival plots with risk tables
- Annotating figures with headers, titles, and footnotes using `gridify`
- Integrating `tflmetaR` to pull annotations from external metadata files

---

## Loading Required Packages

We begin by loading all necessary packages for survival analysis, data manipulation, and figure annotation.

```{r load-packages, message=FALSE, warning=FALSE}
library(tflmetaR)
library(gridify)
library(haven)
library(dplyr)
library(survival)
library(survminer)
```

---

## Loading and Preparing ADTTE Data

The Analysis Dataset for Time-to-Event (ADTTE) follows CDISC ADaM standards. Here we use a synthetic oncology dataset generated by the CDISC Dataset Generator, containing 3 treatment arms and 500 subjects.

```{r load-data}
# get addtte data ----
# synthetic adtte for oncology therapeutic area from CDISC

adtte <- read_xpt(system.file("extdata", "adam_adtte.xpt", package = "tflmetaR"))
```

### Data Transformation

We filter the data for Progression-Free Survival (PFS) analysis and prepare it for survival modeling.

```{r prepare-data}
tte <- adtte |>
  rename(avalc = AVALU) |>
  rename_with(tolower) |>
  filter(paramcd == "PFS") |>
  arrange(desc(trtp)) |>
  relocate(usubjid, trtp, cnsr, evntdesc, avalc, aval, trta) |> #  distinct(usubjid)
  select(usubjid, trtp, cnsr, evntdesc, avalc, aval, trta)

tte$trtp <- factor(tte$trtp)
```

---

## Fitting the Survival Model

We create a survival object and fit a Kaplan-Meier model stratified by treatment group (`trtp`).

```{r fit-survival-model}
# run survival modal ----
surv_object <- Surv(time = tte$aval, event = tte$cnsr)
fit1 <- survfit(surv_object ~ trtp, data = tte)
```

---

## Creating the Survival Plot

The `survminer` package provides `ggsurvplot()` for creating publication-ready Kaplan-Meier curves with integrated risk tables, p-values, and median survival lines.

```{r create-survplot}
# survial plot and table ----
ggsurv <- ggsurvplot(fit1,
  data = tte, risk.table = T, size = 1, # line size
  pval = TRUE,
  pval.size = 4,
  pval.method = TRUE,
  pval.method.size = 3,
  fontsize = 3,
  surv.median.line = "hv",
  tables.theme = theme_survminer() +
    theme(
      axis.text.x = element_text(size = 9, color = "gray30"),
      axis.text.y = element_text(size = 9, color = "gray30"),
      plot.title = element_text(size = 9),
      element_text(size = 9, color = "red")
    )
)
```

---

## Customizing the Risk Table

We modify the risk table labels to display treatment group names in a more readable format.

```{r customize-risk-table}
# risk table -----
t <- ggsurv$table +
  scale_y_discrete(label = c("Treatment 3", "Treatment 2", "Treatment 1"))

p1 <- ggsurvplot(fit1, data = tte, risk.table = TRUE) ## can not add to the gridify directly
```

---

## Combining Plot and Risk Table

Using `ggpubr::ggarrange()`, we combine the survival plot and risk table into a single figure with proper alignment and proportions.

```{r combine-plot-table}
# ggrance the survival plot  and risk table ----
p1 <- ggarrange(
  ggsurv$plot + labs(x = "", y = "Survival Probability") +
    scale_color_discrete() +
    theme(
      axis.title.x = element_text(vjust = 0, size = 9),
      axis.title.y = element_text(vjust = -3, size = 9), # y axis label
      axis.text.y = element_text(size = 9, color = "gray30"), # tick values
      axis.text.x = element_text(size = 9, color = "gray30"),
      legend.title = element_blank()
    ),
  t + labs(y = "") +
    theme(
      axis.title.x = element_text(vjust = 0, size = 9),
      axis.text.x = element_text(size = 9, color = "gray30")
    ),
  heights = c(2, 1.0),
  ncol = 1, nrow = 2, align = "v"
)
```

---

## Annotating with gridify

The `gridify` package provides a framework for adding professional annotations to figures, including headers, titles, footnotes, and page numbers. The `pharma_layout_base()` function creates a layout compliant with pharmaceutical industry standards.

### Basic Gridify Example

This example demonstrates manual annotation using `gridify::set_cell()` to populate each annotation field.

```{r gridify-basic}
## --- use gridify to annotate the figure ---
fig <- gridify(
  p1,
  layout = pharma_layout_base(
    margin = grid::unit(c(1, 1, 1.23, 1), "inches"),
    global_gpar = grid::gpar(fontfamily = "Courier")
  )
) |>
  set_cell("header_left_1", "UCB") |>
  set_cell("header_left_2", "Drug X / Unspecified") |>
  set_cell("header_left_3", "STUDY001") |>
  set_cell("header_right_1", "CONFIDENTIAL") |>
  set_cell("header_right_2", "Final") |>
  set_cell("header_right_3", paste0("Data Cut-Off Date")) |>
  set_cell("output_num", "Figure F01") |>
  set_cell("title_1", "The Kaplan-Meier Curves of Progression-Free Survival Among Treatment Groups") |>
  set_cell("title_2", "Population: Safety Set") |>
  set_cell("title_3", "") |>
  set_cell("note", "The synthetic oncology ADTTE was generated by CDISC Dataset Generater, with 3 treatment arms and 500 subjects.\nNote: The Kaplan-Meier estimate of survival probability at a given time is the product of these conditional\nprobabilities up until that given time. ", mch = 120) |>
  set_cell("footer_left", "Program: f_surv_gridify.R, Source(s): ADTTE") |>
  set_cell("footer_right", paste0("Page ", 1, " of ", 1))

# Optionally export the figure to a PDF file
# gridify::export_to(fig, "f_surv_gridify.pdf")
```

```{r km-plot, dev='png', dpi=300, fig.width=10, fig.height=8, out.width='95%'}
print(fig)
```

---

## Integrating tflmetaR for Metadata Management

The `tflmetaR` package enables **separation of metadata and code** by storing titles, footnotes, and headers in external spreadsheets. This approach:

- Centralizes all annotations in a single location to support audit and traceability
- Reduces code complexity and maintenance burden
- Supports regulatory compliance workflows

### Loading Metadata Files

We read the titles and header information from an Excel spreadsheet.

```{r tflmetar-load-meta}
## --- tflmetaR related code ----

# read headers, titles, and footnotes excel spreadsheet

file_name <- system.file("extdata", "st_titles.xls", package = "tflmetaR")
title_file <- tflmetaR::read_tfile(filename = file_name, sheetname = "Sheet1")
header_file <- tflmetaR::read_tfile(filename = file_name, sheetname = "Headr", validate = FALSE)

fig_number <- "Figure F01"
```

### Extracting Annotations

Using `tflmetaR` accessor functions, we retrieve titles, footnotes, and header content for the specific figure.

```{r tflmetar-extract}
ulheader <- tflmetaR::get_ulheader(header_file)[1, ]
urheader <- tflmetaR::get_urheader(header_file)[1, ]
titles <- tflmetaR::get_title(title_file, tnumber = fig_number)
footnotes <- tflmetaR::get_footnote(title_file, tnumber = fig_number, add_footr_tstamp = FALSE)
pgmname <- tflmetaR::get_pgmname(title_file, tnumber = fig_number)
source <- tflmetaR::get_source(title_file, tnumber = fig_number)

# convert footnote list to string
footnote_str <- paste(unlist(footnotes), collapse = "\n")
```

---

## Creating the Final Annotated Figure

Now we combine `gridify` with `tflmetaR`-extracted metadata to create the final publication-ready figure. This approach ensures that any updates to titles or footnotes in the metadata spreadsheet are automatically reflected in the output.

```{r final-figure}
## --- re-draw graph with pulled titles and footnotes ----
fig2 <- gridify(
  p1,
  layout = pharma_layout_base(
    margin = grid::unit(c(0.5, 1, 0.25, 1), "inches"),
    global_gpar = grid::gpar(fontfamily = "Courier")
  )
) |>
  set_cell("header_left_1", ulheader[[1]]) |>
  set_cell("header_left_2", ulheader[[2]]) |>
  set_cell("header_left_3", ulheader[[3]]) |>
  set_cell("header_right_1", urheader[[1]]) |>
  set_cell("header_right_2", urheader[[2]]) |>
  set_cell("header_right_3", urheader[[3]]) |>
  set_cell("output_num", titles[[1]]) |>
  set_cell("title_1", "") |>
  set_cell("title_2", titles[[2]]) |>
  set_cell("title_3", titles[[3]]) |>
  set_cell("note", footnote_str, mch = 120) |>
  set_cell("footer_left", paste0(
    "Program: ", pgmname, ",  ",
    "Source(s): ", source
  )) |>
  set_cell("footer_right", sprintf("Page %d of %d", 1, 1))

# Optionally export the figure to a PDF file
# gridify::export_to(fig2, "f_surv_tflmetar.pdf")
```

```{r km-plot-tflmetaR, dev='png', dpi=300, fig.width=10, fig.height=8, out.width='95%'}
print(fig2)
```

---

## Summary

This vignette demonstrated a complete workflow for creating annotated Kaplan-Meier survival plots:

1. **Data preparation**: Loading and transforming ADTTE data following CDISC standards
2. **Survival analysis**: Fitting Kaplan-Meier models using the `survival` package
3. **Visualization**: Creating publication-quality plots with `survminer`
4. **Annotation**: Adding professional headers, titles, and footnotes using `gridify`
5. **Metadata integration**: Leveraging `tflmetaR` to externalize annotations for better maintainability

By combining these tools, organizations can produce regulatory-compliant figures while maintaining a clear separation between analysis code and presentation metadata.

---

## Related Resources

### Key Packages Used

| Package | Purpose |
|---------|---------|
| `survival` | Survival analysis and Kaplan-Meier estimation |
| `survminer` | Publication-ready survival plots |
| `gridify` | Professional figure annotation layouts |
| `tflmetaR` | Centralized metadata management |

### tflmetaR Core Functions

| Function | Description |
|----------|-------------|
| `read_tfile()` | Load metadata from Excel files |
| `get_title()` | Extract titles and subtitles |
| `get_footnote()` | Extract footnotes |
| `get_ulheader()` | Extract upper-left header content |
| `get_urheader()` | Extract upper-right header content |

---
