---
title: "Using tflmetaR with ggplot2"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using tflmetaR with ggplot2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(tflmetaR)
library(ggplot2)
library(readxl)
library(patchwork) 
library(gridExtra)
library(grid)
library(magrittr)
library(glue)
library(rstudioapi)
library(dplyr)
```

## 📆 Overview

This vignette demonstrates how to annotate `ggplot2` figures using title and footnote metadata managed via the `tflmetaR` package.

The workflow separates metadata from code by reading header/footer content from an external file, selecting the appropriate entry, and layering it onto a ggplot-based figure.

---

## 📂 Load Metadata

You can load a title/footnote spreadsheet in either CSV or Excel format.

```{r load-data}
# Example: Read from a CSV file
#meta <- read_tfile_csv(system.file("extdata", "demo_titles.csv", package = "tflmetaR"))

# You could also use:
file_name <- file.path("../inst/extdata", "sample_titles2.xlsx")
meta <- read_tfile(filename = file_name, sheetname = "footer")
```

---

## 🎯 Select Metadata for a Specific Plot

```{r select-entry}
# Assume this plot corresponds to TFL number "Figure 1.2"
title_list <- get_title(meta, tnumber = "Figure 1.2")
footnotes <- get_footnote(meta, tnumber = "Figure 1.2")
source_note <- get_source(meta, tnumber = "Figure 1.2")
```

---

## 🎈 Create the ggplot

```{r basic-plot, fig.width= 7, fig.height= 5}
p <- ggplot(mtcars, aes(mpg, wt, color = wt)) +
  geom_point() +
  theme_minimal()
p
```

---

## 📝 Add Title, Footnotes, and Source Using patchwork

```{r final-plot, fig.width= 7, fig.height= 5}
# Combine metadata elements
title_text <- paste(title_list, collapse = "\n")
footnote_text <- paste(unlist(footnotes), collapse = "\n")
source_text <- paste(unlist(source_note), collapse = "\n")

# Construct enhanced figure layout
full_plot <- (
  wrap_elements(grid::textGrob(title_text, gp = grid::gpar(fontsize = 14, fontface = "bold"))) /
  p /
  wrap_elements(grid::textGrob(footnote_text, gp = grid::gpar(fontsize = 9))) /
  wrap_elements(grid::textGrob(source_text, gp = grid::gpar(fontsize = 8, fontface = "italic")))
) +
  plot_layout(heights = c(0.1, 0.8, 0.05, 0.05))

full_plot
```

---

## ✅ Summary

The `tflmetaR` package allows you to:

- Keep titles/footnotes separate from code
- Easily update metadata in one place
- Apply consistent annotations across reports and plots

This approach is ideal for regulatory and clinical settings where version control, traceability, and auditability are essential.

---

## 🔗 See Also

- [`flextable`](https://davidgohel.github.io/flextable/)
- [`patchwork`](https://patchwork.data-imaginist.com/)
- `tflmetaR::get_ulheader()`, `get_urheader()`, `get_timestamp()`

