---
title: "ggplot-example1"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ggplot-example1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


---
title: "Using tfootr with ggplot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using tfootr with ggplot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(tfootr)
library(ggplot2)
library(readxl)
library(patchwork) 
library(gridExtra)
library(grid)
library(magrittr)
library(glue)
library(rstudioapi)
library(dplyr)
```

## 📆 Overview

This vignette demonstrates how to annotate `ggplot2` figures using title and footnote metadata managed via the `tfootr` package.

The workflow separates metadata from code by reading headers/titles/footnotes content from an external file, selecting the appropriate entry, and layering it onto a ggplot-based figure.


## Load Metadata

You can load a title/footnote spreadsheet in either CSV or Excel format.

```{r load-data}
# Example: Read from a CSV file
#m eta <- read_tfile_csv(system.file("../inst/extdata", "sample_titles_in_csv.csv", package = "tfootr"))

# You could also use:
file_name <- file.path("../inst/extdata", "sample_titles2.xlsx")
header_file <- read_tfile(filename = file_name, sheetname = "header")
meta <- read_tfile(filename = file_name, sheetname = "footer")
```


## Select Metadata for a Specific Plot

```{r select-entry}
# Assume this plot corresponds to TFL number "Figure 1.2"
title_list <- get_title(meta, tnumber = "Figure 1.2")
footnotes <- get_footnote(meta, tnumber = "Figure 1.2")
ulheader <- get_ulheader(header_file)
urheader <- get_urheader(header_file)
source_note <- get_source(meta, tnumber = "Figure 1.2")
pgname <- glue("Generated from get.R on ", "\n") #get_refstamp(pname = NULL)
curr_time <- Sys.time()
run_time <- format(curr_time, "%Y-%m-%d %H:%M:%S")
```



## Create the ggplot

A basic ggplot figure without annotations, as shown here:

```{r basic-plot, fig.width= 7, fig.height= 5}
# p <- ggplot(mtcars, aes(mpg, wt)) +
#   geom_point() +
#   theme_minimal()
p <- ggplot(mpg) +
  geom_point(aes(x = displ, y = hwy, color = class))
p
```



## Add Title, Footnotes, and Source Using patchwork

```{r final-plot, fig.width= 7, fig.height= 5}
# Combine metadata elements
title_text <- paste(title_list, collapse = "\n")
footnote_text <- paste(unlist(footnotes), collapse = "\n")
ulheader_text <- paste(unlist(ulheader), collapse = "\n")
urheader_text <- paste(unlist(urheader), collapse = "\n")
source_text <- paste(unlist(source_note), collapse = "\n")
#runtime_text <- paste(run_time, collapse = "\n")

# Construct enhanced figure layout
full_plot <- (
  wrap_elements(grid::textGrob(title_text, gp = grid::gpar(fontsize = 14, fontface = "bold"))) /
  p /
  wrap_elements(grid::textGrob(footnote_text, gp = grid::gpar(fontsize = 9))) /
  wrap_elements(grid::textGrob(source_text, gp = grid::gpar(fontsize = 8, fontface = "italic")))
) +
  plot_layout(heights = c(0.1, 0.8, 0.05, 0.05))

full_plot
```

```{r second-example, fig.width= 7, fig.height= 5}
p <- ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(colour = class)) + 
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE) + 
  theme(legend.position = "right") +
    labs(
    title = {paste(title_list, sep = "\n", collapse = "\n")},
    caption = {paste({paste(footnotes, sep = "\t", collapse = "\n")}, "\n", pgname, run_time, "   Data Source(s): ", source_text)}) +
  theme(plot.caption = element_text(hjust = 0, size = 11)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
p
grid.text(paste(ulheader_text, sep = "\n", collapse = " \n "), x = 0.04, y = 0.95, just = "left",gp = gpar(fontsize = 9, col = "black"))

grid.text(paste(urheader_text, sep = "\n", collapse = " \n "), x = 0.9, y = 0.95, just = "right",gp = gpar(fontsize = 9, col = "black"))
```

## Summary

The `tfootr` package allows you to:

- Keep titles/footnotes separate from code
- Easily update metadata in one place
- Apply consistent annotations across reports and plots

This approach is ideal for regulatory and clinical settings where version control, traceability, and auditability are essential.



## 🔗 See Also

- [`flextable`](https://davidgohel.github.io/flextable/)
- [`patchwork`](https://patchwork.data-imaginist.com/)
- `tfootr::get_ulheader()`, `get_urheader()`, `get_timestamp()`

