---
title: "Using tflmetaR with gt for Professional Tables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using tflmetaR with gt for Professional Tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
# Load required packages
library(tflmetaR)
library(gt)
library(dplyr)
```

## Introduction

In clinical research and regulatory reporting, maintaining consistency and accuracy in table annotations—including titles, subtitles, footnotes, and data sources—is critical. Traditional approaches often embed these annotations directly within analysis scripts, leading to maintenance challenges, inconsistencies across outputs, and increased risk of errors during updates.

The `tflmetaR` package addresses these challenges by implementing a **separation of metadata and code** design pattern. By externalizing table annotations into structured spreadsheets (Excel or CSV), `tflmetaR` enables:

- **Centralized metadata management**: All titles, footnotes, and source information are stored in a single, auditable location
- **Reduced code complexity**: Analysis scripts remain focused on data processing and visualization
- **Enhanced reproducibility**: Metadata changes do not require modifications to validated code
- **Regulatory compliance**: Supports audit-friendly workflows common in pharmaceutical and clinical settings

This vignette demonstrates how to integrate `tflmetaR` with the `gt` package to create publication-quality tables with externally managed annotations.

---

## Loading Table Metadata

The first step in using `tflmetaR` is loading the metadata file that contains your table annotations. The package supports both Excel (`.xlsx`, `.xls`) and CSV formats through dedicated functions:

- `read_tfile()`: Reads metadata from Excel files with optional sheet specification
- `read_tfile_csv()`: Reads metadata from CSV files

The metadata file should contain standardized columns including `PGMNAME` (program name), `TTL1` (primary title), `FOOT1` (first footnote), and `SOURCE` (data source). Additional columns for subtitles (`TTL2`, `TTL3`, ...), additional footnotes (`FOOT2`, `FOOT3`, ...), population definitions, and bylines are also supported.

```{r load-data}
# Load the sample metadata file included with the package
file_path <- system.file("extdata", "sample_titles2.xlsx", package = "tflmetaR")

# Read the metadata from the 'footer' sheet
meta <- read_tfile(filename = file_path, sheetname = "footer")

# Preview the structure of the metadata
cat("Metadata columns:", paste(names(meta), collapse = ", "))
```

The metadata file serves as a centralized repository for all table annotations, enabling consistent application across multiple outputs and facilitating updates without code modifications.

---

## Extracting Table Annotations

Once metadata is loaded, `tflmetaR` provides a suite of accessor functions to retrieve specific annotation components for a given table. Tables can be identified by their TFL number (`tnumber`) or program name (`pname`).

### Core Accessor Functions

| Function | Description |
|----------|-------------|
| `get_title()` | Retrieves title and subtitle fields (TTL1, TTL2, ...) along with population |
| `get_footnote()` | Retrieves footnote fields (FOOT1, FOOT2, ...) with optional timestamp |
| `get_source()` | Retrieves data source information |
| `get_pop()` | Retrieves population definition |
| `get_byline()` | Retrieves byline fields (BYLINE1, BYLINE2, ...) |
| `get_pgmname()` | Retrieves the program name |

```{r select-entry}
# Extract metadata components for a specific table
# Tables are identified by their TFL number (e.g., "Table 2.1")
title_info <- get_title(meta, tnumber = "Table 2.1")
footnotes <- get_footnote(meta, tnumber = "Table 2.1", add_footr_tstamp = FALSE)
source_info <- get_source(meta, tnumber = "Table 2.1")
population <- get_pop(meta, tnumber = "Table 2.1")
bylines <- get_byline(meta, tnumber = "Table 2.1")
pgm_name <- get_pgmname(meta, tnumber = "Table 2.1")

# Display extracted metadata
cat("Title:", unlist(title_info$TTL1), "\n")
cat("Population:", unlist(population), "\n")
cat("Byline:", unlist(bylines), "\n")
cat("Program:", unlist(pgm_name), "\n")
```

Each function returns a named list containing the non-missing fields, allowing flexible integration with various table rendering packages.

---

## 📄 Create and Annotate a Table

```{r basic-table}
# Create a simple table
data <- head(mtcars, 5)

# Build title text and footnote text
title_text <- paste(title_list, collapse = "\n")
footnote_text <- paste(unlist(footnotes), collapse = "\n")
source_text <- paste(unlist(source_note), collapse = "\n")

# Render kable table with annotations
kable(data, format = "html", caption = title_text) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  add_footnote(label = footnote_text, notation = "symbol") %>%
  add_footnote(label = source_text, notation = "none")
```

---

## ✅ Summary

With the `tflmetaR` package, you can:

- Manage annotation metadata externally
- Easily reuse footnotes and titles
- Apply consistent formatting to `kable` tables across reports

This is especially valuable for clinical and regulatory reports that must adhere to strict documentation standards.

---

## 🔗 Related Packages

- [`kableExtra`](https://haozhu233.github.io/kableExtra/)
- [`flextable`](https://davidgohel.github.io/flextable/)
- `tflmetaR::get_title()`, `get_footnote()`, `get_source()`

---


