---
title: "Using tflmetaR with gt for Professional Tables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using tflmetaR with gt for Professional Tables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
# Load required packages
library(tflmetaR)
library(gt)
library(dplyr)
```

## Introduction

In clinical research and regulatory reporting, maintaining consistency and accuracy in table annotations—including titles, subtitles, footnotes, and data sources—is critical. Traditional approaches often embed these annotations directly within analysis scripts, leading to maintenance challenges, inconsistencies across outputs, and increased risk of errors during updates.

The `tflmetaR` package addresses these challenges by implementing a **separation of metadata and code** design pattern. By externalizing table annotations into structured spreadsheets (Excel or CSV), `tflmetaR` enables:

- **Centralized metadata management**: All titles, footnotes, and source information are stored in a single, auditable location
- **Reduced code complexity**: Analysis scripts remain focused on data processing and visualization
- **Enhanced reproducibility**: Metadata changes do not require modifications to validated code
- **Regulatory compliance**: Supports audit-friendly workflows common in pharmaceutical and clinical settings

This vignette demonstrates how to integrate `tflmetaR` with the `gt` package to create publication-quality tables with externally managed annotations.

---

## 📂 Load Metadata

You can load a title/footnote spreadsheet from CSV or Excel format.

```{r load-data}
# Example CSV file from the package
# meta <- read_tfile_csv(system.file("../inst/extdata", "sample_titles_in_csv.csv", package = "tftr"))

# You could also use:
file_name <- system.file("extdata", "sample_titles2.xlsx", package = "tflmetaR")
meta <- read_tfile(filename = file_name, sheetname = "footer")
```

---

## 🎯 Select Entry for a Specific Table

```{r select-entry}
# Select the metadata entry for a table
title_list <- get_title(meta, tnumber = "Table 2.1")
footnotes <- get_footnote(meta, tnumber = "Table 2.1")
source_note <- get_source(meta, tnumber = "Table 2.1")
```

---

## 📄 Create and Annotate a Table

```{r basic-table}
# Create a simple table
data <- head(mtcars, 5)

# Build title text and footnote text
title_text <- paste(title_list, collapse = "\n")
footnote_text <- paste(unlist(footnotes), collapse = "\n")
source_text <- paste(unlist(source_note), collapse = "\n")

# Render kable table with annotations
kable(data, format = "html", caption = title_text) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  add_footnote(label = footnote_text, notation = "symbol") %>%
  add_footnote(label = source_text, notation = "none")
```

---

## ✅ Summary

With the `tflmetaR` package, you can:

- Manage annotation metadata externally
- Easily reuse footnotes and titles
- Apply consistent formatting to `kable` tables across reports

This is especially valuable for clinical and regulatory reports that must adhere to strict documentation standards.

---

## 🔗 Related Packages

- [`kableExtra`](https://haozhu233.github.io/kableExtra/)
- [`flextable`](https://davidgohel.github.io/flextable/)
- `tflmetaR::get_title()`, `get_footnote()`, `get_source()`

---


