---
title: "table-example"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{table-example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

---
title: "Using tfootr with kableExtra"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using tfootr with kableExtra}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(TFootr)
library(kableExtra)
library(dplyr)
```

## 📆 Overview

This vignette demonstrates how to annotate tables created with `knitr::kable()` and styled using `kableExtra`, using title and footnote metadata managed via the `tfootr` package.

---

## 📂 Load Metadata

You can load a title/footnote spreadsheet from CSV or Excel format.

```{r load-data}
# Example CSV file from the package
#meta <- read_tfile_csv(system.file("../inst/extdata", "sample_titles_in_csv.csv", package = "TFootr"))

# You could also use:
file_name <- file.path("../inst/extdata", "sample_titles2.xlsx")
meta <- read_tfile(filename = file_name, sheetname = "footer")
```

---

## 🎯 Select Entry for a Specific Table

```{r select-entry}
# Select the metadata entry for a table
title_list <- get_title(meta, tnumber = "Table 2.1")
footnotes <- get_footnote(meta, tnumber = "Table 2.1")
source_note <- get_source(meta, tnumber = "Table 2.1")
```

---

## 📄 Create and Annotate a Table

```{r basic-table}
# Create a simple table
data <- head(mtcars, 5)

# Build title text and footnote text
title_text <- paste(title_list, collapse = "\n")
footnote_text <- paste(unlist(footnotes), collapse = "\n")
source_text <- paste(unlist(source_note), collapse = "\n")

# Render kable table with annotations
kable(data, format = "html", caption = title_text) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  add_footnote(label = footnote_text, notation = "symbol") %>%
  add_footnote(label = source_text, notation = "none")
```

---

## ✅ Summary

With the `tfootr` package, you can:

- Manage annotation metadata externally
- Easily reuse footnotes and titles
- Apply consistent formatting to `kable` tables across reports

This is especially valuable for clinical and regulatory reports that must adhere to strict documentation standards.

---

## 🔗 Related Packages

- [`kableExtra`](https://haozhu233.github.io/kableExtra/)
- [`flextable`](https://davidgohel.github.io/flextable/)
- `tfootr::get_title()`, `get_footnote()`, `get_source()`

---


