---
knit: function(inputFile, encoding){
      out_dir <- "/Biocdm/Products/MT1621/iseema/gsp/analysis1a/tfls";
      rmarkdown::render(inputFile,
                        encoding=encoding,
                        output_file=paste(out_dir, "Figure_testa.pdf", sep="/")) }
output: 
  pdf_document:
    latex_engine: xelatex


mainfont: "Courier New"
sansfont: "Courier New"
monofont: "Courier New"

classoption: landscape
geometry: left=0.5in, right=0.5in, top=0.8in, bottom=0.8in, footskip=0.2in


header-includes:
 - \usepackage{xcolor}
 - \usepackage{sectsty}
 - \usepackage{fancyhdr}
 - \usepackage{lastpage}
 - \usepackage{datetime2}
---
\allsectionsfont{\normalfont\fontsize{9}{10}\selectfont\center}
\fontsize{9}{10}
\selectfont
\pagestyle{fancy}
\setlength\headheight{46.5pt}
\setlength\headsep{0pt}
\fancyhead[L]{\fontsize{9}{10} \selectfont UCB \\ MT1621 Response to Questions \\~}
\fancyhead[R]{\fontsize{9}{10} \selectfont Version: FINAL \\ Confidential \\~}
\fancyfoot[L]{\fontsize{9}{10} \selectfont Generated from g\_frst\_hr\_u.Rmd on `r toupper(format(Sys.time(), "%d%b%Y %H:%M:%S"))`. Source Dataset(s): ADTTE}
\fancyfoot[C]{\fontsize{9}{10} \selectfont }
\fancyfoot[R]{\fontsize{9}{10} \selectfont Page \thepage\ of \pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\raggedright


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = T, warning = T, message = T)
knitr::opts_chunk$set(fig.width=10, fig.height=5.6) 
Sys.setenv(PATH='/opt/software/texlive/2019/bin/x86_64-linux:/usr/lib/jvm/java-1.8.0-openjdk/bin:/opt/rh/devtoolset-3/root/usr/bin:/opt/rh/devtoolset-7/root/usr/bin:/usr/local/bin:/usr/bin:/usr/local/sbin:/usr/sbin')
```


```{r loadpackages, include=FALSE}
library(tidyr)
library(dplyr)
library(stringr)
library(ggplot2)
library(haven)
library(patchwork)
```
##Figure 16.13.1.1.1 Treated Subjects from Clinical Trials (101, 102 and 107) and Modified Untreated Subjects : Forest Plots of Hazard Ratios ??? Matched-Pairs Cox models with and without Age of TK2d Symptom Onset Continuous as Covariate and Marginal Cox Models - Time to Death from TK2d Symptom Onset and from Treatment
##EP - All Treated and Modified Untreated Subjects
```{r}
library(flextable)

# Sample data frame
df <- data.frame(
  Species = as.factor(c("setosa", "versicolor", "virginica")),
  Sepal.Length_mean = c(5.006, 5.936, 6.588),
  Sepal.Length_sd = c(0.35249, 0.51617, 0.63588),
  Sepal.Width_mean = c(3.428, 2.77, 2.974),
  Sepal.Width_sd = c(0.37906, 0.3138, 0.3225)
)

# Create a flextable object
ft <- flextable(df)

# Apply separate_header() to split headers
#ft <- flextable::separate_header(ft, opts = c("span-top", "bottom-vspan"))

# Display the flextable

```
