---
title: "use_ggplot"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{use_ggplot}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=10, 
  fig.height=6
)
```


## Introduction


This document is designed to provide you with practical examples of how to use `headr` to add headers and footnotes to your figure outputs. 

Here's an example of using `headr` with `ggplot` to generate figure outputs:

```{r setup}
# Load the necessary libraries
library(headr)
library(ggplot2)
#library(tables)
#library(mtcars)
library(magrittr)
library(glue)
library(dplyr)
# (to use |> version 4.1.0 of R is required, for lower versions we recommend %>% from magrittr)
 

fname <- "C:/Users/u066227/Rproj/sample_titles.xls"
# hfooter_file <- read_footer("C:/Users/u066227/Rproj/sample_titles.xls")
# hfooter_list <- select_with_number(df =hfooter_file,  tnumber = "Table 2.1")
# 
# head_list <- ht_list %>% select(starts_with("TTL"), POPULATION) 
# head_list <- Filter(function(x) !is.na(x), head_list)
# foot_list <- ht_list %>% select(starts_with("FOOT")) 

head_list <- header_for_flextable(filename = fname, tnumber = "Figure 1.1")
foot_list <- footer_for_flextable(filename = fname, tnumber = "Figure 1.1")

fig <- ggplot(mpg, aes(displ, hwy)) + 
  geom_point(aes(colour = class)) + 
  geom_smooth(method = "lm", se = FALSE) + 
  theme(legend.position = "right") +
    labs(
    title = {paste(head_list, sep = "\n", collapse = "\n ")},
    caption = {paste(foot_list, sep = "\t", collapse = "\n ")}) +
      theme(plot.caption = element_text(hjust = 0, size = 11)) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")
  )

fig

```

## Examples

### Simple Examples
#### Simple Layout with margins

In these examples, we're creating a simple plot using the `mtcars` dataset and adding headers and footnotes to it.

Here, we're creating a line plot of `mpg` against `displ` from the `mtcars` dataset. 



#### `show(layout)` method:
